SHIV: Reducing Supervisor Burden using Support Vectors to Efficiently Learn Robot Control Policies from Demonstrations

### 2015-08-22 paper.tex
Anca and Florian's feedback

\begin{figure}[t!]
\centering
\includegraphics[width=\columnwidth]{figures/teaser.pdf}
\caption{ 
Illustrations of the different examples we evaluated our algorithm, SHIV, on. A) Driving simulator where the goal is to learn a controller on HOG features extracted from synthetic images to keep a car on a polygonal track. B) Grasping in clutter experiment in a physics simulator, where the goal is to learn a controller from human demonstrations to grasp the green object while not moving the red square objects off the gray boundary representing a table. C) A experiment on the Da Vince Research Kit to learn a controller from a surgeon's demonstrations of correcting a needle before suturing in surgery. 
}
\vspace*{-10pt}
\label{fig:dis_traveled}
\end{figure}

@@

\section{Related Work}
\adnote{stick to either learning from demonstration or imitation learning}
Below we summarize related work in robot\adnote{robot learnign too broad, active approaches to learning from demonstration perhaps} learning and in risk evaluation.

@@

Stream-based active learning has two ingredients: a stream, given here by the policy, and a query selection method. Typical query selection methods are estimator-centric: they evaluate the estimator's confidence, e.g. distance from the classification hyperplane \cite{tong2002support}, implicitly assuming a stationary distribution --- that the new data is sampled from the same distribution as the previous training data. Although such methods have been proposed for online imitation learning (see dogged learning \cite{chernova2009interactive,grollman2007dogged}),
online imitation learning violates the stationary distribution assumption because each new policy induces a new distribution of states. An alternative to distance-based measures of confidence is query by committee, which uses a committee of hypothesized classifiers\adnote{don't have to be classifiers, repalce with estimators maybe} that are trained on different subset of training data and only labels points for which they disagree\adnote{you evaluate confidence based on disagreement, not only label points for which they disagree; so rephrase to "..subsets of data. Confidence is based on the level of agreement or disagreement among these hypotheses."}. Judah et al. has proposed using this approach towards online imitation learning\cite{judah2011active,judah2012active}, however it has been shown when training and test distributions are different this can perform worse than randomly selecting points\adnote{points confusing; suggest: than randomly selecting which states to query} \cite{burbidge2007active}.

In SHIV, we use a query selection method that allows a non-stationary distribution, evaluating confidence based on novelty (is the state an outlier to the training distribution)
and on historically mislabeled regions. %Our approach builds on the One Class SVM method for approximating quantile level sets \cite{}, which has already been used in novelty detection on high-dimensional image data \cite{}. 
Our results suggest that this method is a better predictor of misclassification risk than a typical, estimator-centric query selector (either distance or query by committee based).

@@

\begin{figure*}[ht]
\centering

\caption{An illustrative example of different active learning techniques. The shaded and empty circles represent
different labeled classes $0$ and $1$. The red empty circles represent label $1$, but are encountered on a different
distribution then what was trained on
{\color{blue} FP:can we make this more formal sounding? Represent samples from a different distribution?} 
As illustrated, Confidence and Query by Committee methods
measure
{\color{blue} FP:estimate} states far away from the decision boundary of the estimator as having low risk,
however novelty detection methods measure {\color{blue} FP:estimate} the level of risk as
{\color{blue}FP:by} how far a point is from the trained data. Our
modified risk method attempts to measure risk both as a function of how far it
{\color{blue}it gramatically refers to risk,
but you mean sample "it$\mapsto$a sample"?} is from data trained on
{\color{blue}grammatically strange$\mapsto$from the training data} and the decision boundary of the estimator. 
}
\label{fig:support_example}
\end{figure*}

\section{Problem Statement}
\subsection{Notation and Background}
We denote by $\mathcal{X}$ the set consisting of \emph{observable states} for a robot task, consisting, for example of
high-dimensional vectors corresponding to images from a camera, or robot joint angles and object poses in the environment.
We furthermore consider a set $\mathcal{U}$ of \emph{allowable control inputs} for the robot, which can be discrete or
continuous. We assume that Markovian dynamics, such that the the probability of state $\mathbf{x_{t+1}}\in
\mathcal{X}$ can be determined from the previous state $\mathbf{x}_t\in\mathcal{X}$ and control input $\mathbf{u}_t\in
\mathcal{U}$. Under this assumption $p(\bx_{N+1}|\bu_{N},\bx_{N}, \ldots, \bu_{0}, \bx_{0})=p(\bx_{N+1}|\bu_{N}, \bx_N)$.  
We denote the probability density over the initial state also by $p:\mathcal{X}\to
\mathbb{R}$. By a trajectory, we mean a finite series of pairs of $T$ states visited and corresponding
control inputs at these states $\tau = (\mathbf{x}_0,\mathbf{u}_0, ...., \mathbf{x}_T,\mathbf{u}_T)$, where $\bx_i\in \mathcal{X}$
and $\bu_i\in \mathcal{U}$ for $i\in \{0, \ldots, T\}$ and some $T\in \mathbb{N}$.  
For a given trajectory $\tau$ as above, we denote by $\hat{\tau}$ the corresponding trajectory in state space,
$\hat{\tau} = (\bx_0,....,\bx_T)$ with $\bx_i\in\mathcal{X}$ for $i\in \{0, \ldots, T\}$.

A policy is a function $\pi: \mathcal{X} \to \mathcal{U}$ from states to control inputs. 
We consider a space of policies $\pi_{\theta}:\mathcal{X}\to \mathcal{U}$ parameterized by a
parameter $\theta\in \mathbb{R}^d$. Any such policy $\pi_{\theta}$ in an environment with probabilistic initial state density and Markovian dynamics
induces a density on state-space trajectories of length $T+1$: $p(\tau | \theta)=
p(\bx_0)\prod_{i=0}^{T-1}p(\bx_{t+1}|\pi_{\theta}(\bx_t),\bx_t)$, for $\tau = (\bx_0, \bx_1, \ldots, \bx_N)\in
\mathcal{X}^{N+1}$.

Let $p(\bx_t|\theta)$ denote the value of the density of states visited at time $t$ if the robot follows the policy
$\pi_{\theta}$ for $t-1$ time steps. This density can be computed by marginalization: $p(\bx_t|\theta) =
\int_{\bx_{t-1}}...\int_{\bx_1} p((\bx_t,...,\bx_1)|\theta) d\bx_{t-1}...d\bx_1$. Following \cite{ross2010reduction},
the average density on states is now defined as $p(\bx|\theta) = \frac{1}{T} \sum^T_{t=1} p(\bx_t|\theta)$.
While we do not assume analytic knowledge of the distributions corresponding to: $p(\bx_{t+1}|\bx_t,\bu_t)$, $p(\bx_0)$, $p(\bx_{\bx}|
\bx)$ or $p(\bx|\theta)$, we assume that we have a stochastic real robot or a simulator such that for any state
$\bx_t$ and control $\bu_t$, we can sample the $\bx_{t+1}$ from the density $p(\bx_{t+1}|\pi_{\theta}(\bx_t),\bx_t)$. 
Therefore, when 'rolling out' trajectories under a policy
$\pi_{\theta}$, we utilize the robot or a simulator to sample the resulting stochastic trajectories rather than
estimating $p(\bx|\theta)$ itself.

Typically, the objective of policy learning is to find a policy that minimizes some known cost function
$C(\tau) = \sum^T_{t=1} c(\bx_t,\bu_t)$ of a trajectory $\tau = (\mathbf{x}_0,\mathbf{u}_1, ...., \mathbf{x}_T,\mathbf{u}_T)\in (\mathcal{X}\times
\mathcal{U})^{T+1}$. The cost function $c:\mathcal{X}\times \mathcal{U}\to \mathbb{R}$ is typically user defined and task specific. 
For example, in task of inserting a peg into a hole, the distance between the peg's current and desired final state can
be considered \cite{levine2015end}.  
In our learning from demonstration setup, the robot does not have access to the cost function itself. Instead, it only has access to a
a `supervisor' that we assume uses some near-optimal policy $\tilde{\pi}$ to minimize $C(\tau)$. We are furthermore given
an initial set of $N$ stochastic demonstration trajectories $\mathcal{D} = \lbrace \tilde{\tau}^1,...,\tilde{\tau}^N \rbrace$. 
which are the result of the supervisor applying this policy.

Next, we define a `surrogate' loss function $l:\mathcal{U}\times \mathcal{U}\to \mathbb{R}$, which provides a distance
measure between any pair of control values. In the continuous case, we consider $l(\bu_0,\bu_1) = ||\bu_0-\bu_1||^2$,
while in the discrete case $l(\bu_0,\bu_1) = 1$ if $\bu_0 \neq \bu_1$ and $l(\bu_0, \bu_1)=0$ otherwise.

Given a candidate policy $\pi_{\theta}$, we then use the surrogate loss function to approximately measure how "close" the policy's
returned control intput $\pi_{\theta}(\bx)\in \mathcal{U}$ at a given state $\bx\in \mathcal{X}$ is to the supervisor's policy's control output
$\tilde{\pi}(\bx)\in \mathcal{U}$. Our objective then consists of determining a policy $\pi_{\theta}$ minimizing the expected surrogate loss where the expectation is taken over the distribution of states.

 \vspace{-2ex}
\begin{align}\label{eq:LFD_obj}
\underset{\theta}{\min} \: E_{p(\bx|\theta)} [l(\pi_\theta(\bx),\tilde{\pi}(\bx))].
\end{align}

The optimization problem in Eq. \ref{eq:LFD_obj} is  typically non-convex. It can also be difficult to solve, since we only have access to samples from the supervisor's policy $\tilde{\pi}$ arising from the demonstration trajectories $\mathcal{D}$. To address these issues Stephane and Ross proposed an iterative method known as DAgger \cite{ross2010reduction}.

 \subsection{DAgger: Dataset Aggregation}
Note that Eq. \ref{eq:LFD_obj} involves both the parameterized policy $\pi_{\theta}$ and the average distribution of
states with respect to the policy, $p(\bx|\theta)$. Let us recall that Dataset Aggregation (DAgger) is an iterative algorithm developed by \cite{ross2010reduction} which iterates two basic steps for $K$ iterations to find an optimal policy under this objective. 
DAgger iteratively minimizes $\pi_{\theta}$ on the collected dataset of examples $\mathcal{D}$ and then adds new states
of high probability under $\pi_\theta$ by sampling via policy `roll outs' from $p(\bx|\theta)$. 
\fpnote{DAgger and \acro~formal algorithms side by side would be better, also the Step 1/2 below can be
made more precise then and refer to the algorithm}

\subsubsection{Step 1}
\fpnote{this is really start of iteration `i', as we circle back}
At iteration $k=0$ an initial dataset $\mathcal{D}_0$ of $N_0$ supervisor demonstration trajectories 
with $T+1$ time steps: $\tau^i=(\bx_0^i, \bu_0^i, \ldots, \bx_T^i, \bu_T^i)\in (\mathcal{X}\times\mathcal{U})^{T+1}$ for $i\in \{0, \ldots, N_0\}$
mapping states to control
inputs is provided. DAgger solves the following optimization problem with respect to the surrogate loss function:

 \vspace{-2ex}
\begin{align}\label{eq:super_objj}
\theta_{k} = \underset{\theta}{\argmin} \: \sum_{i=1}^{N_0}\sum_{t=0}^T l(\pi_{\theta}(\bx_t^{i}),\bu_{t}^i).
\end{align}

This problem can be considered as a supervised learning problem which can be solved using standard machine
learning techniques, for example a support vector machine or a neural network.

To handle the fact that the supervisor's policy can be noisy, a zero-mean noise term $\epsilon$ 
can be considered as present in policy's output. We employ a regularization technique in the optimization which is
used to control the smoothness of the function that is fit to the sampled data. In practice this regularization corresponds to a penalty term on either the L2 norm on the weights for regression based techniques or the slack coefficient for support vector machines \cite{scholkopf2002learning}.

 \subsubsection{Step 2}
DAgger rolls out the policy $\pi_{\theta_{k=1}}$ to sample states that are likely to occur under $\theta_{k=1}$. For every state visited, DAgger requests the supervisor to provide the appropriate control/label. Formally, for a given sampled trajectory  $\tau = (\bx_0,\bu_0,...,\bx_T,\bu_T )$. The supervisor provides labels $\tilde{\bu}_t$, where $\bu_t \sim \pi(\bx_t) + \epsilon$ for $t\in \{0, \ldots, T\}$.
The states and labeled controls are then aggregated into the next data set of demonstrations $\mathcal{D}_1$. 
\fpnote{terms such as aggregated etc would be much clearer if we just refer to a particular line in the algorithm}

Steps 1 and 2 are repeated for $K$ iterations or until 
the cumulative surrogate is smaller than some predefined threshold\footnote{In the orginal DAgger the policy rolled out
was stochastically mixed with the supervisor, thus with probability $\beta$ it would either take the supervisor's action
or the robots. The use of this stochasitically mix policy was for theoretical analysis. In practice, it is recommended
to set $\beta = 0$ to avoid biasing the sampling~\cite{NIPS2014_5421,ross2010reduction}}.

Dagger has been experimentally shown to perform  well in both simulation and real world experiments and furthermore under the assumption of the policy being no regret has theoretical properties of being linearly in time horizon T to the supervisor's policy. 
\fpnote{How about: Dagger has been experimentally shown to perform well in both simulation and real world
    experiments. Under a technical `no regret' assumption, Ross et al.\cite{ross2013learning} furthermore
deduced error bounds on the expected loss which scale linearly in the time horizon $T$.}

However, policy rollouts can in practice result in the robot entering dangerous states. For example, when DAgger was used on a
quad-copter the robot repeatedly crashed into trees before successfully learning to avoid them \cite{ross2013learning}.
We work under the assumption that a state can be risky for 2 reasons: 1) it lies in an area with a low density of
previously trained states, which can cause our policy to mis-predict the supervisor and incur high surrogate
loss \cite{tokdar2010importance} 2) the surrogate loss of the current policy at the current state is high, so that the state, although visited, does
not model the supervisor's control inputs correctly. Furthermore providing correct control inputs or ``labeling'' at each
iteration and each state in DAgger can be tedious or costly.

To address these limitations, we propose to 1) avoid highly risky states entirely by early termination of the policy
roll-outs that enter these regions and 2) to recognize non-risky states and to not request supervisor input for these,
thereby reducing the burden on the human trainer.

One way to do this is to define risk in terms of distance to a nearest neighbor, but this approach is vulnerable to
outliers in the known set of states. Another option is to fit a density to the known set of states, which requires an
exponential number of samples in the dimension of the state space \cite{nadaraya1964estimating}. We propose using a
modified version of the technique known as the One Class SVM that implicitly approximately estimates a boundary of a user defined quantile of
a density representing the training data in $\mathcal{X}$ \cite{scholkopf2001estimating}.
\fpnote{Refer to figure, replace "one way to do"  by "an approach to this problem is to" or similar - less informal}

\subsection{Estimation of Quantile Level Sets}\label{sec:level}
\fpnote{my git commit on this section was overwritten by an old commit at version 10 - reintroduced modifications here,
Michael, please check this is still accurate with recent modifications}
We consider the problem of estimating the quantile level-sets of a distribution $P$ on a set $\mathcal{X}$ by means of a finite set of
independent and identically distributed samples $\mathbf{x}_1,...,\mathbf{x}_n\in \mathcal{X}$.
In most general terms, the quantile function for $P$ and subject to a class of measurable subsets $\mathcal{G}$ of $\mathcal{X}$ is
defined by
\begin{align}\label{eq:quantile}
U(\gamma) = \mbox{inf} \lbrace \lambda(G):P(G) \geq \gamma, G \in \mathcal{G} \rbrace \: 0<\gamma \leq 1
\end{align} 
$\lambda:\mathcal{G}\to \mathbb{R}$ above denotes a volume measure which most commonly is given by the Lebesgue measure.
Suppose furthermore that $G:[0,1]\to \mathcal{G}$ assigns a set $G(\gamma) \in \mathcal{G}$ that attains the infinum
measure (i.e. volume) for each $\gamma\in [0,1]$ (this set is in general not necessarily unique). 
$G(\gamma)$ denotes a set of minimum measure $G \in \mathcal{G}$ with $P(G(\gamma))\ge \gamma$. Note in particular that $G(1)$ is the support of the density $p$ corresponding to $P$, if $p$ exists.

>>>>

\begin{figure}[t!]
\centering
\includegraphics[width=\columnwidth]{figures/teaser.pdf}
\caption{ \adnote{make this a 2 by 2 grid; a b c d, where d is the result; make the result graph more square}
For a given budget of human supervision, SHIV improves performance (reduced policy cost) in our three test domains: 1) a driving simulator where the goal is to learn a controller on HOG features extracted from synthetic images to keep a car on a polygonal track; 2) push-grasping in clutter in a physics simulator, where the goal is to learn a controller from human demonstrations to grasp the green object while not letting the red square objects fall off the gray boundary representing a table; 3) learning a controller for the Da Vinci Research Kit from a surgeon's demonstrations of correcting a needle before suturing in surgery. 
}
\vspace*{-10pt}
\label{fig:dis_traveled}
\end{figure}

\begin{figure*}[t]
\centering

\caption{\adnote{A comparison of different query selection strategies for active learning on a non-stationary distribution. The shaded and empty circles are training data from two classes, $0$ and $1$ respectively. The red empty circles are samples from a new distribution (produced by executing the learned policy), and belong to class $1$. Typical strategies assign high risk to states close the decision boundary (a), or for which a set of estimators disagree (b). Neither of these apply to our new samples in red. In contrast, we use a strategy that is amenable to non-stationary distributions by assigning high risk to novel states (c) and states in historically mislabeled regions. }
%Confidence and Query by Committee methods measure
%{\color{blue} FP:estimate} states far away from the decision boundary of the estimator as having low risk,
%however novelty detection methods measure {\color{blue} FP:estimate} the level of risk as
%{\color{blue}FP:by} how far a point is from the trained data. Our
%modified risk method attempts to measure risk both as a function of how far it
%{\color{blue}it gramatically refers to risk,
%but you mean sample "it$\mapsto$a sample"?} is from data trained on
%{\color{blue}grammatically strange$\mapsto$from the training data} and the decision boundary of the estimator. 
}
\label{fig:support_example}
\end{figure*}

@@

\section{Related Work}
Below we summarize related work in active approaches to robot learning from demonstration, as well as in risk evaluation techniques that we build upon.

@@

Stream-based active learning has two ingredients: a stream, given here by executing the current best policy, and a query selection method deciding whether or not to query. Typical query selection methods are estimator-centric: they evaluate the estimator's confidence, e.g. distance from the classification hyperplane \cite{tong2002support}, as in Fig. \ref{fig:support_example}(a). An alternative to distance-based measures of confidence is query by committee \adnote{cite}, which uses a committee of hypothesized estimators that are trained on different subsets of the training data. Confidence here is based on the level of agreement or disagreement among these hypotheses, with higher levels of disagreement for a state leading to higher chances of querying that state (Fig. \ref{fig:support_example}(b)).

Both approaches implicitly assume a stationary state distribution --- that the new data is sampled from the same distribution as the previous training data. Although such methods have been previously proposed for online imitation learning (see \cite{chernova2009interactive,grollman2007dogged} for the former and \cite{judah2011active,judah2012active} for the latter),
online imitation learning violates the stationary distribution assumption because each new policy induces a new distribution of states. This can have negative consequences for learning: it has been shown that when training and test distributions are different, query by committee can perform worse than randomly selecting  which states to query \cite{burbidge2007active}.\adnote{merged the two types since they have the same problem, and references the figure}

In SHIV, we use a query selection method that allows for a non-stationary distribution, evaluating confidence based on \emph{novelty} (is the state an outlier to the training distribution), as in Fig. \ref{fig:support_example}(c). Additionally, we also assign lower confidence to  historically \emph{mislabeled} regions, as in Fig. \ref{fig:support_example}(d). %Our approach builds on the One Class SVM method for approximating quantile level sets \cite{}, which has already been used in novelty detection on high-dimensional image data \cite{}. 
Our results in Sec. \ref{}\adnote{fill in} suggest that this method is a better predictor of misclassification risk than a typical, estimator-centric query selector (be it distance or query by committee based).

@@

\section{Problem Statement}
\noindent\textbf{Policies and State Densities.}
We denote by $\mathcal{X}$ the set consisting of \emph{observable states} for a robot task, consisting, for example, of
high-dimensional vectors corresponding to images from a camera, or robot joint angles and object poses in the environment.
We furthermore consider a set $\mathcal{U}$ of \emph{allowable control inputs} for the robot, which can be discrete or
continuous. We assume Markovian dynamics, such that the probability of state $\mathbf{x_{t+1}}\in
\mathcal{X}$ can be determined from the previous state $\mathbf{x}_t\in\mathcal{X}$ and control input $\mathbf{u}_t\in
\mathcal{U}$. Thus, we assume that 
$$p(\bx_{t+1}|\bu_{t},\bx_{t}, \ldots, \bu_{0}, \bx_{0})=p(\bx_{t+1}|\bu_{t}, \bx_t)$$
We assume a probability density over initial states $p(\bx_0)$.
%We denote the probability density over the initial state also by $p:\mathcal{X}\to \mathbb{R}$.

A trajectory $\hat{\tau}$ is a finite series of $T+1$ pairs of states visited and corresponding
control inputs at these states, $\hat{\tau} = (\mathbf{x}_0,\mathbf{u}_0, ...., \mathbf{x}_T,\mathbf{u}_T)$, where $\bx_t\in \mathcal{X}$
and $\bu_t\in \mathcal{U}$ for $t\in \{0, \ldots, T\}$ and some $T\in \mathbb{N}$.  
For a given trajectory $\hat{\tau}$ as above, we denote by ${\tau}$ the corresponding trajectory in state space,
${\tau} = (\bx_0,....,\bx_T)$.

A policy is a function $\pi: \mathcal{X} \to \mathcal{U}$ from states to control inputs. 
We consider a space of policies $\pi_{\theta}:\mathcal{X}\to \mathcal{U}$ parameterized by some $\theta\in \mathbb{R}^d$. Any such policy $\pi_{\theta}$ in an environment with probabilistic initial state density and Markovian dynamics
induces a density on state-space trajectories of length $T+1$: $$p(\tau | \theta)=
p(\bx_0)\prod_{i=0}^{T-1}p(\bx_{t+1}|\pi_{\theta}(\bx_t),\bx_t)$$

Let $p(\bx_t|\theta)$ denote the value of the density of states visited at time $t$ if the robot follows the policy
$\pi_{\theta}$ from time $0$ to time $t-1$.\adnote{shouldnt you then define this above insted of the probability of a trajectory?} This density can be computed by marginalization: $p(\bx_t|\theta) =
\int_{\bx_{t-1}}...\int_{\bx_1} p((\bx_t,...,\bx_1)|\theta) d\bx_{t-1}...d\bx_1$. Following \cite{ross2010reduction}, we can compute
the average density on states for \emph{any} timepoint by 
\begin{equation}
p(\bx|\theta) = \frac{1}{T} \sum^T_{t=1} p(\bx_t|\theta)
\label{eq:density}
\end{equation}

While we do not assume analytic knowledge of the distributions corresponding to: $p(\bx_{t+1}|\bx_t,\bu_t)$, $p(\bx_0)$, $p(\bx_{\bx}|
\bx)$\adnote{this is a typo} or $p(\bx|\theta)$, we assume that we have a stochastic real robot or a simulator such that for any state
$\bx_t$ and control $\bu_t$, we can sample the $\bx_{t+1}$ from the density $p(\bx_{t+1}|\pi_{\theta}(\bx_t),\bx_t)$. 
Therefore, when 'rolling out' trajectories under a policy
$\pi_{\theta}$, we utilize the robot or a simulator to sample the resulting stochastic trajectories rather than
estimating $p(\bx|\theta)$ itself.

\noindent\textbf{Objective.} Typically, the objective of policy learning is to find a policy that minimizes some known cost function
$C(\tau) = \sum^T_{t=1} c(\bx_t,\bu_t)$ of a trajectory $\tau$. The cost function $c:\mathcal{X}\times \mathcal{U}\to \mathbb{R}$ is typically user defined and task specific. 
For example, in task of inserting a peg into a hole, the distance between the peg's current and desired final state can
be considered \cite{levine2015end}.  
In our learning from demonstration setup, the robot does not have access to the cost function itself. Instead, it only has access to 
a `supervisor' that we assume uses some near-optimal policy $\tilde{\pi}$ to minimize $C(\tau)$. We are furthermore given
an initial set of $N$ stochastic demonstration trajectories $\lbrace \tilde{\tau}^1,...,\tilde{\tau}^N \rbrace$. 
which are the result of the supervisor applying this policy. This induces a training dataset $\mathcal{D}$ of all state-control input pairs from the demonstrated trajectories. The goal is to produce a low cost policy for the task based on these demonstrations.

We define a `surrogate' loss function $l:\mathcal{U}\times \mathcal{U}\to \mathbb{R}$, which provides a distance
measure between any pair of control values. In the continuous case, we consider $l(\bu_0,\bu_1) = ||\bu_0-\bu_1||^2$,
while in the discrete case $l(\bu_0,\bu_1) = 1$ if $\bu_0 \neq \bu_1$ and $l(\bu_0, \bu_1)=0$ otherwise.

Given a candidate policy $\pi_{\theta}$, we then use the surrogate loss function to approximately measure how "close" the policy's
returned control input $\pi_{\theta}(\bx)\in \mathcal{U}$ at a given state $\bx\in \mathcal{X}$ is to the supervisor's policy's control output
$\tilde{\pi}(\bx)\in \mathcal{U}$.

Following \cite{ross2010reduction}, our objective then consists of determining a policy $\pi_{\theta}$ minimizing the expected surrogate loss, where the expectation is taken over the distribution of states induced by the policy across any time point in the horizon:

 \vspace{-2ex}
\begin{align}\label{eq:LFD_obj}
\underset{\theta}{\min} \: E_{p(\bx|\theta)} [l(\pi_\theta(\bx),\tilde{\pi}(\bx))].
\end{align}
with $p(\bx|\theta)$ from Eq. \ref{eq:density}.

If the robot could learn the policy perfectly, this state density would match the one encountered in the user examples. But any small error will make the state distribution deviate from the training one, and this the motivation for an iterative algorithm like DAgger and now SHIV.

% The optimization problem in Eq. \ref{eq:LFD_obj} is  typically non-convex. It can also be difficult to solve, since we only have access to samples from the supervisor's policy $\tilde{\pi}$ arising from the demonstration trajectories $\mathcal{D}$. To address these issues Stephane and Ross proposed an iterative method known as DAgger \cite{ross2010reduction}.

\section{SHIV}
\adnote{putting shiv and dagger side by side. also, notice how I am setting up the rest of the section by explaining what will follow; also, avoid filler sentences like "note that ..." when you are not following that up with a point; avoid filler words like two "basic" steps -- it's just steps, why are they basic?}

Both SHIV and DAgger \cite{ross2010reduction} solve the minimization in Eq. \ref{eq:LFD_obj} by iterating two steps: 1) find an optimal $\theta$ using the training data $\mathcal{D}$ thus far, and 2) execute the policy induced by the current $\theta$, and ask for labels for the encountered states. However, instead of querying the supervisor for every new state, SHIV will actively decide whether the state is risky enough to warrant a query, thereby reducing supervisory burden.

%  \subsection{DAgger: Dataset Aggregation}
% Note that Eq. \ref{eq:LFD_obj} involves both the parameterized policy $\pi_{\theta}$ and the average distribution of
% states with respect to the policy, $p(\bx|\theta)$. Let us recall that Dataset Aggregation (DAgger) is an iterative algorithm developed by \cite{ross2010reduction} which iterates two basic steps for $K$ iterations to find an optimal policy under this objective. 
%DAgger iteratively minimizes $\pi_{\theta}$ on the collected dataset of examples $\mathcal{D}$ and then adds new states
%of high probability under $\pi_\theta$ by sampling via policy `roll outs' from $p(\bx|\theta)$. 
%\fpnote{DAgger and \acro~formal algorithms side by side would be better, also the Step 1/2 below can be
%made more precise then and refer to the algorithm}

\subsection{Step 1.}
The first step of any iteration $k$ is to find a $\theta_k$ that minimizes surrogate loss on the current dataset $\mathcal{D_k}=\{(x_i,u_i)|i\in\{1,\ldots,M\}\}$ of demonstrated state-control pairs (initially just the set $D$ from above):
% \fpnote{this is really start of iteration `i', as we circle back}
% At iteration $k=0$ an initial dataset $\mathcal{D}_0$ of $N_0$ supervisor demonstration trajectories 
% with $T+1$ time steps: $\tau^i=(\bx_0^i, \bu_0^i, \ldots, \bx_T^i, \bu_T^i)\in (\mathcal{X}\times\mathcal{U})^{T+1}$ for $i\in \{0, \ldots, N_0\}$
% mapping states to control
% inputs is provided. DAgger solves the following optimization problem with respect to the surrogate loss function:

 \vspace{-2ex}
\begin{align}\label{eq:super_objj}
\theta_{k} = \underset{\theta}{\argmin} \: \sum_{i=1}^{M} l(\pi_{\theta}(\bx_i),\bu_i).
\end{align}

\adnote{simplified D above to just be pairs so that shiv makes sense too, as it doesnt add full trajectories}
This problem can be treated as a supervised learning problem which can be solved using standard machine
learning techniques, for example a support vector machine or a neural network.

To handle the fact that the supervisor's policy can be noisy, a zero-mean noise term $\epsilon$ 
can be considered as present in policy's output. We employ a regularization technique in the optimization which is
used to control the smoothness of the function that is fit to the sampled data. In practice this regularization corresponds to a penalty term on either the L2 norm on the weights for regression based techniques or the slack coefficient for support vector machines \cite{scholkopf2002learning}.\adnote{is this just us or also dagger?}

 \subsection{Step 2}
The second step starts in boht SHIV and DAgger by rolling out the policy $\pi_{\theta_{k}}$ to sample states that are likely to occur under $\theta_{k}$.

What happens next, however, differs. For every state visited, DAgger requests the supervisor to provide the appropriate control/label. Formally, for a given sampled trajectory  $\hat{\tau} = (\bx_0,\bu_0,...,\bx_T,\bu_T )$, the supervisor provides labels $\tilde{\bu}_t$, where $\tilde{\bu}_t \sim \pi(\bx_t) + \epsilon$ for $t\in \{0, \ldots, T\}$.
The states and labeled controls are then aggregated into the next data set of demonstrations $\mathcal{D}_{k+1}$:
$$D_{k+1}=\mathcal{D}_k \cup \{(\bx_t,\tilde{\bu_t})\|t\in\{0,\ldots,T\}\} $$
%\fpnote{terms such as aggregated etc would be much clearer if we just refer to a particular line in the algorithm}

Providing correct control inputs or ``labeling'' at each
iteration and each state in DAgger can be tedious, imposing a large burden on the supervisor. \adnote{add in the discussion that another advantage of measuring risk is the early termination which can avoid disaster}
Instead of asking the supervisor for labels at all visited states $\{\bx_0,..,\bx_T\}$, SHIV uses a measure of \emph{risk} to actively \emph{decide} whether a label is necessary. For every state encountered $\bx_t$, it applies a \emph{decision function} $g_{\sigma}$, which we introduce in the next section, parametrized by a risk-sensitivity parameter $\sigma$ which will capture how smooth we expect the policy function to be: smoother functions imply that training examples have larger support, and data nearby will be less risky.

SHIV only asks for supervision on states for which $g(\bx)=-1$: \adnote{don't know what a good way to mention the termination condition is, maybe just a footnote or in discussion at the end as an advantage}
$$D_{k+1}=\mathcal{D}_k \cup \{(\bx_t,\tilde{\bu_t})\|t\in\{0,\ldots,T\},g(\bx_t)=-1\}$$

Steps 1 and 2 are repeated for $K$ iterations or until 
the cumulative surrogate is smaller than some predefined threshold\footnote{In the original DAgger the policy rolled out
was stochastically mixed with the supervisor, thus with probability $\beta$ it would either take the supervisor's action
or the robots. The use of this stochastically mix policy was for theoretical analysis. In practice, it is recommended
to set $\beta = 0$ to avoid biasing the sampling~\cite{NIPS2014_5421,ross2010reduction}}.

Dagger has been experimentally shown to perform well in both simulation and real world
    experiments\cite{}\adnote{cite}. Under a `no regret' assumption, Ross et al.\cite{ross2013learning} furthermore
deduced error bounds on the expected loss, which scale linearly in the time horizon $T$ --- as opposed to offline methods which scale quadratically. For a conservative threshold $\rho$ on risk, the SHIV performance is analogous to DAgger without asking the expert when it is absolutely not necessary. However, if $\rho$ is too high, then SHIV will not collect new that that is needed and revert back to the performance of offline methods. We analyze the dependence on $\rho$ in our experimental section. Our results suggest that SHIV is robust to this parameter selection. \adnote{must add this experiment!}

\subsection{Measuring Risk}

\adnote{somewhere in this section refer to the second figure}

% However, policy rollouts can in practice result in the robot entering dangerous states. For example, when DAgger was used on a
% quad-copter, the robot repeatedly crashed into trees before successfully learning to avoid them \cite{ross2013learning}.
We work under the assumption that a state can be risky for 2 reasons: 1) it lies in an area with a low density of
previously trained states, which can cause our policy to mis-predict the supervisor and incur high surrogate
loss \cite{tokdar2010importance}, or 2) the surrogate loss of the current policy at the current state is high, so that the state, although visited, does
not model the supervisor's control inputs correctly.

% To address these limitations, we propose to 1) avoid highly risky states entirely by early termination of the policy
% roll-outs that enter these regions and 2) to recognize non-risky states and to not request supervisor input for these,
% thereby reducing the burden on the human trainer.

%One way to do this is to define risk in terms of distance to a nearest neighbor, but this approach is vulnerable to
%outliers in the known set of states. Another option is to fit a density to the known set of states, which requires an
%exponential number of samples in the dimension of the state space \cite{nadaraya1964estimating}.\adnote{this stuff is in the related work too}

To evaluate risk in high-dimensional state spaces (such as HOG features on images, like in our driving simulator testcase), we use
modified version of the technique known as the One Class SVM that implicitly approximately estimates a boundary of a user defined quantile of
a density representing the training data in $\mathcal{X}$ \cite{scholkopf2001estimating}.
%\fpnote{Refer to figure, replace "one way to do"  by "an approach to this problem is to" or similar - less informal}

%\subsection{Estimation of Quantile Level Sets}\label{sec:level}
\fpnote{my git commit on this section was overwritten by an old commit at version 10 - reintroduced modifications here,
Michael, please check this is still accurate with recent modifications}
We consider the problem of estimating the quantile level-sets of a distribution $P$ on a set $\mathcal{X}$ by means of a finite set of
independent and identically distributed samples $\mathbf{x}_1,...,\mathbf{x}_n\in \mathcal{X}$.
In most general terms, the quantile function for $P$ and subject to a class of measurable subsets $\mathcal{G}$ of $\mathcal{X}$ is
defined by
\begin{align}\label{eq:quantile}
U(\gamma) = \mbox{inf} \lbrace \lambda(G):P(G) \geq \gamma, G \in \mathcal{G} \rbrace \: 0<\gamma \leq 1
\end{align} 
$\lambda:\mathcal{G}\to \mathbb{R}$ above denotes a volume measure which most commonly is given by the Lebesgue measure.
Suppose furthermore that $G:[0,1]\to \mathcal{G}$ assigns a set $G(\gamma) \in \mathcal{G}$ that attains the infinum
measure (i.e. volume) for each $\gamma\in [0,1]$ (this set is in general not necessarily unique). 
$G(\gamma)$ denotes a set of minimum measure $G \in \mathcal{G}$ with $P(G(\gamma))\ge \gamma$. Note in particular that $G(1)$ is the support of the density $p$ corresponding to $P$, if $p$ exists.
